---
title: "project.Rmd"
author: "Yijun Jiang"
date: "4/21/2017"
output: html_document
---
### Read in Training Data
```{r}
library(MASS)
library(dplyr)
```

To get started read in the training data:
```{r read-data}
load("ames_train.Rdata")
```
## Part I: Simple Model

In the first model you are allowed only limited manipulations of the original data set to predict the sales price `price`. You are allowed to take power transformations of the original variables [square roots, logs, inverses, squares, etc.] but you are NOT allowed to create interaction variables. This means that a variable may only be used once in an equation [if you use $ x^2$ don’t use $x$]. Additionally, you may eliminate any data points you deem unfit. This model should have a minimum r-square of 73% (in the original units) and contain at least 6 variables but fewer than 20.   

Deal with missing value
```{r}
apply(ames_train,2,function(x) sum(is.na(x)))
```

```{r}
#for (i in 1: ncol(ames_train)){
  #if (any(is.na(ames_train[,i][[1]])) & is.factor(ames_train[,i][[1]])){
    ames_train[,i][[1]] = addNA( ames_train[,i][[1]])
  #}else if (any(is.na(ames_train[,i][[1]])) & (!is.factor(ames_train[,i][[1]]))){
    # ames_train[,i][[1]][is.na(ames_train[,i][[1]])] = mean( ames_train[,i][[1]][!is.na(ames_train[,i][[1]])])
  #}
#}
```

```{r}
suppressMessages(library(gbm))
boost.car =gbm(price ~ .,
data=ames_train,
n.trees =5000, interaction.depth =4)
```
```{r}
rmse = function(y, ypred) {
  rmse = sqrt(mean((y - ypred)^2))
  return(rmse)
}
```

```{r}
results = summary(boost.car,plotit = FALSE)
results
```

```{r model1}
model1 = lm(log(price)~  Overall.Qual + Neighborhood+area+BsmtFin.SF.1 + Total.Bsmt.SF+Bsmt.Qual +Garage.Area
            + Kitchen.Qual + Mas.Vnr.Area + Lot.Area + Year.Remod.Add + Overall.Cond+Year.Built
            +Lot.Frontage+Central.Air+Exterior.1st+BsmtFin.SF.2+Wood.Deck.SF+MS.Zoning,data =ames_train)
summary(model1)
```
### Model Evaluation on Test Data
Create predicted values for price using your model using the testing data

```{r predict-model1, echo=FALSE}
#Yhat = predict(model1, newdata=ames_test, interval = "pred")
```


```{r}
model1 =gbm(price~.,data =ames_train,n.trees =5000, interaction.depth =15)
```


```{r}
predictions = as.data.frame(predict(model1, newdata=ames_test, interval = "pred",n.trees = 5000))
predictions$PID = ames_test$PID
save(predictions, file="predict.Rdata")
```
