---
title: "writeup"
author: "BayeStar"
date: "April 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

### Exploratory data analysis

```{r}
load("ames_train.Rdata")
```

Exploratory data analysis (20 points): must include three correctly labeled graphs and an explanation that highlight the most important features that went into your model building.

* Figure 1 shows the boxplot of prices by each neighborhood. It is apparent that prices vary a lot by different neighborhoods. Outliers are also captured and we considered delete some of them. 
* Figure 2 is a scatterplot of prices against porch area. The left panel corresponds to houses without porch, while the right with porch. We can see that the relationship differs between with and without porch, so we decided to add a dummy variable indicating the existence of porch and an interaction between the dummy and porch area.
* Figure 3 shows the prices of houses built in various years. There is a non-linear trend, so we considered adding quadratic term of year to capture the non-linearity. 
* Figure 4 depicts the relationship between prices and total square by neighborhood. (Only four of neighborhoods are shown for plot purpose.) Prices and total square are linearly related, but the linear relationship changes across neigborhood. We considered adding interaction term of total square and neighborhood to capture different slopes. 
* Figure 5 is a vilion plot showing the distribution of prices by overall quality. Prices are higher, but also vary more for higher quality.  
```{r}
# price v.s. neighborhood
ggplot(ames_train, aes(x=reorder(Neighborhood, price, FUN=median), y=price))+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))+
  geom_boxplot()+
  xlab('')+
  ylab('Price')+
  ggtitle("Figure 1. Prices by Neighborhood")

# price v.s. Enclosed.Porch
ames_train %>%
  mutate(Porch = ifelse(Enclosed.Porch==0, "w/ porch", "w/o porch")) %>%
  ggplot(aes(x=Enclosed.Porch, y=price))+
  geom_point(col="dark blue", cex=1, alpha=0.5)+
  facet_wrap(~Porch)+
  xlab("Porch Area")+
  ylab("Price")+
  theme_bw()+
  ggtitle("Figure 2. Prices against Porch Area")

# year
ggplot(ames_train, aes(x=Year.Built, y=price,alpha = 0.5))+
  geom_point(col="dark blue", cex=1)+
  geom_smooth(method = "lm", formula = y ~ x + I(x^2)+I(x^3), 
              col="black", se=F, size=0.7)+
  theme_bw()+
  theme(legend.position="none")+
  ylab('Price')+
  xlab('')+
  ggtitle("Figure 3. Prices of Houses Built in Years")
  
# Neighborhood and TotalSq
ames_train %>%
  filter(Neighborhood=="OldTown" |Neighborhood=="NridgHt"|
           Neighborhood=="IDOTRR"|Neighborhood=="Veenker") %>%
  ggplot(aes(x=TotalSq, y=price))+
  geom_point(col="dark blue", cex=1, alpha=0.5)+
  facet_wrap(~Neighborhood)+
  geom_smooth(method="lm", se=F, col="black", size=0.7)+
  theme_bw()+
  xlab("Total Square")+
  ylab("Price")+
  ggtitle("Figure 4. Prices by Total Square in Different Neighborhoods")

# Overall.Qual
ggplot(ames_train, aes(x=factor(Overall.Qual), y=price))+
  geom_violin(aes(color=factor(Overall.Qual), 
                  fill = factor(Overall.Qual), alpha=0.5))+
  theme_bw() +
  theme(legend.position="none")+
  xlab("Overal Quality")+
  ylab("Price")+
  ggtitle("Figure 5. Prices by Overall Quality")
```

### Simple Model 
2. Development and assessment of an initial model from Part I (10 points)

Preprocess - clean data + remove outliers
```{r}
# clean data
clean_data = function(xdata){
xdata %>%
    mutate(# replace NAs with new levels
           Alley = as.factor(ifelse(is.na(as.character(Alley)), 
                                    "No alley access", as.character(Alley))),
           Bsmt.Qual = as.factor(ifelse(as.character(Bsmt.Qual)=="Po", 
                                           "Fa", as.character(Bsmt.Qual))),
           Bsmt.Qual = as.factor(ifelse(is.na(as.character(Bsmt.Qual)), 
                                           "No Basement", as.character(Bsmt.Qual))),
           Bsmt.Cond = as.factor(ifelse(is.na(as.character(Bsmt.Cond)), 
                                           "No Basement", as.character(Bsmt.Cond))),
           BsmtFin.Type.1 = as.factor(ifelse(is.na(as.character(BsmtFin.Type.1)), 
                                           "No Basement", as.character(BsmtFin.Type.1))),
           BsmtFin.Type.2 = as.factor(ifelse(is.na(as.character(BsmtFin.Type.2)), 
                                           "No Basement", as.character(BsmtFin.Type.2))),
           Bsmt.Exposure = as.factor(ifelse(is.na(as.character(Bsmt.Exposure))|
                                              as.character(Bsmt.Exposure) == "", 
                                           "No Basement", as.character(Bsmt.Exposure))),
           Bsmt.Unf.Rate.SF = ifelse(Total.Bsmt.SF!=0, Bsmt.Unf.SF/Total.Bsmt.SF, 0),
           Bsmt.Full.Bath = ifelse(is.na(Bsmt.Full.Bath),0,Bsmt.Full.Bath),
           Bsmt.Half.Bath = ifelse(is.na(Bsmt.Half.Bath),0,Bsmt.Half.Bath),
           Fireplace.Qu = as.factor(ifelse(is.na(as.character(Fireplace.Qu)), 
                                           "No Fireplace", as.character(Fireplace.Qu))),
           Garage.Type = as.factor(ifelse(is.na(as.character(Garage.Type)),
                                           "No Garage", as.character(Garage.Type))),
           Garage.Finish = as.factor(ifelse(is.na(as.character(Garage.Finish))|
                                              as.character(Garage.Finish) == "",
                                           "No Garage", as.character(Garage.Finish))),
           Garage.Qual = as.factor(ifelse(as.character(Garage.Qual)=="Ex", 
                                          "Gd", as.character(Garage.Qual))),
           Garage.Qual = as.factor(ifelse(is.na(as.character(Garage.Qual)),
                                           "No Garage", as.character(Garage.Qual))),
           Garage.Cond = as.factor(ifelse(as.character(Garage.Cond)=="Ex", 
                                          "Gd", as.character(Garage.Cond))),
           Garage.Cond = as.factor(ifelse(is.na(as.character(Garage.Cond))|as.character(Garage.Cond)=="Po",
                                           "No Garage", as.character(Garage.Cond))),
           # deal with new level issue in test data
           Fence = as.factor(ifelse(is.na(as.character(Fence)),
                                           "No Fence", as.character(Fence))),
           Misc.Feature = as.factor(ifelse(is.na(as.character(Misc.Feature)),
                                           "None", as.character(Misc.Feature))),
           Mas.Vnr.Type = as.factor(ifelse(as.character(Mas.Vnr.Type) == "",
                                           "None", as.character(Mas.Vnr.Type))),
           Mas.Vnr.Area = ifelse(is.na(Mas.Vnr.Area),0,Mas.Vnr.Area),
           Kitchen.Qual = as.factor(ifelse(as.character(Kitchen.Qual)=="Po", 
                                           "Fa", as.character(Kitchen.Qual))),
           Heating.QC = as.factor(ifelse(as.character(Heating.QC)=="Po", 
                                         "Fa", as.character(Heating.QC))),
           Electrical = as.factor(ifelse(as.character(Electrical) == "", 
                                         "SBrkr", as.character(Electrical))),
           Condition.2 = as.factor(ifelse(as.character(Condition.2) %in%
                                            c("Artery","RRAn","RRAe"),
                                          "Feedr", as.character(Condition.2))),
           Neighborhood = as.factor(ifelse(as.character(Neighborhood)=="Blueste",
                                          "NPkVill", as.character(Neighborhood))),
           # create new variables
           Enclosed.Porch.is = as.factor(ifelse(Enclosed.Porch==0,"N","Y")),
           Pool.Area = as.factor(ifelse(Pool.Area==0,"N", "Y")),
           Garage.Yr.Blt = ifelse(is.na(Garage.Yr.Blt), Year.Built-2, Garage.Yr.Blt)
          )%>%
    dplyr::select(-c(Lot.Frontage,Pool.QC,Pool.Area))
}

# remove outliers
ames_train = clean_data(ames_train)
ames_train = ames_train[ames_train$price<580000,-1]
remove_idx1 = c(1:nrow(ames_train))[ames_train$Neighborhood %in%c("Gilbert")&ames_train$price>350000]
remove_idx2 = c(1:nrow(ames_train))[ames_train$Neighborhood %in%c("NAmes")&ames_train$price>300000]
remove_idx3 = c(1:nrow(ames_train))[ames_train$Neighborhood %in%c("Landmrk","GrnHill")]
ames_train = ames_train[-c(remove_idx1, remove_idx2, remove_idx3),]
```

* Initial model: must include a summary table and an explanation/discussion for variable selection. Interpretation of coefficients desirable for full points.

We first putted in a few variables that are intuitively important as our base model and kept all the variables put in significant. The variables we chose include `area`, `Year.Built`, `Garage.Area`, `Overall.Qual`, `Kitchen.Qual`, `Garage.Cond`, etc. Based on this base model, forward selection was used to select more variables that can be included in our model and improve the prediction accuracy. In this process we try to avoid the variables might be correlated with other variables, such as TotalSq an area. Finally we end up with a model with 20 variables.$\\\\$

Based on the summary table of the selected model, all the selected continuous variables are extremely significant. These continuous variables include `area`, `Year.Built`, `Year.Remod.Add`, `Garage.Area`, `Overall.Qual`, `Lot.Area`, `BsmtFin.SF.1`, `Overall.Cond`, `Total.Bsmt.SF`, `Bsmt.Full.Bath` and `Screen.Porch`. Additionally, some categorical variables having a lot of significant levels are selected, such as `Neighborhood`, `Kitchen.Qual`, `Exter.Qual`.



```{r}
model1 = lm(price^(0.25) ~ area + Year.Built + Year.Remod.Add + Garage.Area + Overall.Qual +  log(Lot.Area) + BsmtFin.SF.1 + Overall.Cond + Total.Bsmt.SF + Central.Air + Bsmt.Full.Bath + Screen.Porch + Kitchen.Qual + Exter.Qual + Bldg.Type + Bsmt.Qual + Garage.Cond + Neighborhood + Heating.QC, data=ames_train)
summary(model1)
```


* Model selection: must include a discussion

* Residual: must include a residual plot and a discussion

* RMSE: must include an RMSE and an explanation  (other criteria desirable)

* Model testing: must include an explanation

### Complex Model

### Conclusion